
---
title: "Descriptive analyses"
#bibliography: references.bib
author: "Bas Hofstra"
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r format(Sys.time(), '%B, %Y')`

<br>

----

This is the code with which we render our descriptive analyses

<br>

----

# Initatiating R environment

Start out with a custom function to load a set of required packages.
  
```{r pack, eval=TRUE}
# packages and read data
rm(list = ls())

# (c) Jochem Tolsma
fpackage.check <- function(packages) {
  lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE, repos = "http://cran.us.r-project.org")
      library(x, character.only = TRUE)
    }
  })
}
packages = c("haven", "coda", "matrixStats", "parallel", "MASS", "doParallel", "dplyr", "cowplot", 
             "tidyverse", "naniar", "dotwhisker" ,"gt", "reshape2", "VGAM", "expss", "Hmisc",
             "poweRlaw", "fitdistrplus", "grid", "ggplotify", "smplot2")
fpackage.check(packages)
rm(packages)
load("data/dutch_netsize_analyses_revision_2.rda")
```
<br>

----

# Descriptives independent variables

The code below generates the descriptive values which are used for Tables 1 and 2.
  

```{r descind, eval = TRUE}

# descriptive table not automated: TABLE 1
table(df$work)
psych::describe(df$work)
psych::describe(df$hhsize)

table(df$agecat)


# 1 Maj
# 2 west
# 3 nonwest


table(df$income)
# 1 < modal
# 2 > modal
# 3 unknown

psych::describe(df$worthhouse)


table(df$woman)
psych::describe(df$woman)



table(df$opl2)
summary(df$opl2)
# 1 prim/sec
# 2 lower tert
# 3 higher tert

#psych::describe(df$neighdens)


# this is for the correlation table
#df$educ3 <- df$opl
#df$educ3 <- as.numeric(as.character(df$educ3))

table(df$migr3)

#cor(as.matrix(df[df$income!=3,c("income","worthhouse")]))

#rcorr(as.matrix(df[,c("work", "hhsize", "leeftijd10", "woman", "educ3", "neighdens", "worthhouse" )]), type = "pearson")

```

<br>

```{r tablehomogeneity, eval = TRUE}
# table 2
summary(df$perwomen)
summary(df[df$woman == 1, c("perwomen")])
summary(df[df$woman == 0, c("perwomen")])

summary(df$permen)
summary(df[df$woman == 1, c("permen")])
summary(df[df$woman == 0, c("permen")])

summary(df$samegender)
summary(df[df$woman == 1, c("samegender")])
summary(df[df$woman == 0, c("samegender")])

summary(df$pereduchigh)
summary(df[df$opl2 == 1, c("pereduchigh")])
summary(df[df$opl2 == 0, c("pereduchigh")])

summary(df$pereduclow)
summary(df[df$opl2 == 1, c("pereduclow")])
summary(df[df$opl2 == 0, c("pereduclow")])

summary(df$sameeduc)
summary(df[df$opl2 == 1, c("sameeduc")])
summary(df[df$opl2 == 0, c("sameeduc")])

summary(df$netsover6)

```

----

# Comparing naive estimations

The code below generates Figure 1 from the paper and runs a number of correlations found on page X.
  

```{r naicomps, eval = TRUE}
# 
# comps <- gather(df[, c("nsize_naive", "nsize_b1b2", "nsize_b3", "nsize_n1", "nsize_n2", "nsize_n3", "nsize_n123")])
# 
# comparisons <- ggplot(comps[comps$value < 2000, ], aes(x=value, color = key)) + 
#   geom_density(size = 0.5) + 
#   theme_minimal() +
#   labs(x = "Network size estimate", y = "Density") +
#   scale_color_manual(name = "Estimation scenario", 
#                       breaks = c("nsize_naive", "nsize_b1b2", "nsize_b3", "nsize_n1", "nsize_n2", "nsize_n3", "nsize_n123"), 
#                        labels = c("Naive estimator: all X's", 
#                                   "Battery 1: Tert. educ., elec vehicle, etc.", 
#                                   "Battery 2: all names", 
#                                   "Names 1: Sophie, Anna, Thomas, Willem", 
#                                   "Names 2: Julia, Elisabeth, Max, Ali",
#                                   "Names 3: Sanne, Cornelia, Kevin, Mohammed",
#                                   "Names1+2+3"),
#                      values = c("#000000", "#E69F00", "#56B4E9", "#009E73", 
#                           "#0072B2", "#D55E00", "#CC79A7", "darkgrey"))
# 
# # save
# ggsave("output/comparisons.pdf", plot = comparisons, device = "pdf",
#        scale = 1, width = 10, height = 5.5, units = c("in"),
#        dpi = "retina")
# 
# 
# 
# # # correlation battery 1 and 2
# # cor(df$nsize_b1, df$nsize_b2)
# # 
# # # correlation battery 1 and all
# # cor(df$nsize_b1, df$nsize_naive)
# 
# # correlation battery 1+2 and all
# cor(df$nsize_b1b2, df$nsize_naive)
# 
# # correlation battery 3 (names) and all
# cor(df$nsize_b3, df$nsize_naive)
# 
# # correlation names 1 and all
# cor(df$nsize_n1, df$nsize_naive)
# 
# # correlation names 2 and all
# cor(df$nsize_n2, df$nsize_naive)
# 
# # correlation names 3 and all
# cor(df$nsize_n3, df$nsize_naive)
# 
# # correlation names 123 and all
# cor(df$nsize_n123, df$nsize_naive)
# 
# # sample 5 columns of scenarios and correlate with naive estimand
# set.seed(1987)
# sample(13:183, 5)
# 
# # correlation scenario 85 and naive
# cor(df$nsize_naive, df[, 85])
# 
# # correlation scenario 36 and naive
# cor(df$nsize_naive, df[, 36])
# 
# # correlation scenario 150 and naive
# cor(df$nsize_naive, df[, 150])
# 
# # correlation scenario 14 and naive
# cor(df$nsize_naive, df[, 14])
# 
# # correlation scenario 125 and naive
# cor(df$nsize_naive, df[, 125])
# 
# comparisons

```




<br>

----

# Comparing Bayesian estimations

The code below generates Figure 2 from the paper associated with the text on page X.
  
```{r bayecomps, eval = TRUE}
# 
# netsizes <- read.table(file = "data/dutch_netsize_desc.txt")
# 
# # VIZ of netsize
# netsize_l <- gather(netsizes)
# dens1 <- ggplot(netsize_l[netsize_l$value < 2000,], aes(x=value, color = key)) + 
#   geom_density(alpha = .2, size = 0.1) + 
#   theme_minimal() + 
#   theme(legend.position = "none") +
#   labs(x = "Network size estimate", y = "Density") +
#   geom_vline(xintercept=as.numeric(psych::describe(netsize_l[, 2])[3]), color = "darkgrey", linetype = 2) +
#   geom_vline(xintercept=as.numeric(psych::describe(netsize_l[, 2])[5]), color = "darkgrey", linetype = 2) +
#   annotate("text", x = 490, y = 0.0004, color = "darkgrey", angle = 90, 
#            label = paste0("Mean = ", round(as.numeric(psych::describe(netsize_l[, 2])[3]), digits = 0))) +
#   annotate("text", x = 340, y = 0.0004, color = "darkgrey", angle = 90, 
#            label = paste0("Median = ", round(as.numeric(psych::describe(netsize_l[, 2])[5]), digits = 0))) +
#   ggtitle("B) Distribution of network sizes, all scenarios")
# 
# 
# # Get lower triangle of the correlation matrix
# get_lower_tri<-function(cormat){
#   cormat[upper.tri(cormat)] <- NA
#   return(cormat)
# }
# mat <- cor(netsizes)
# lower_tri <- get_lower_tri(mat)
# melted_cormat <- reshape2::melt(lower_tri) # data.table also has melt funciton that won't work on matrices
# melted_cormat$Var2 <- as.character(melted_cormat$Var2)
# melted_cormat$Var1 <- as.character(melted_cormat$Var1)
# melted_cormat <- melted_cormat[!melted_cormat$Var2 == melted_cormat$Var1,]
# melted_cormat <- melted_cormat[!is.na(melted_cormat$value), ]
# 
# # Viz of correlations between netsize estimates
# dens2 <- ggplot(melted_cormat, aes(x=value)) + 
#   geom_density(size = 0.3) + 
#   xlim(0.915, 1) + 
#   theme_minimal() + 
#   labs(x="Pearson correlation", y = "Density") +
#   geom_vline(xintercept = .95, color = "darkgrey", linetype = 2) +
#   ggtitle("A) Distributions of correlations between network size scenarios")  +
#   geom_segment(x = 0.95, y = 125, xend = .99, yend = 125, linetype = 2, color = "darkgrey",
#                 arrow = arrow(length = unit(0.25, "cm")))  +
#   annotate("text", x = .97, y = 135, color = "darkgrey", angle = 0, 
#            label = paste0(round(nrow(melted_cormat[melted_cormat$value > .95, ])/nrow(melted_cormat)*100, digits = 1),"% of correlations are > .95"))
# 
# # lay 'm out on the grid           
# denses <- plot_grid(dens2, dens1, nrow = 1) 
# 
# # save
# ggsave("output/densities.pdf", plot = denses, device = "pdf",
#        scale = 1, width = 12, height = 4, units = c("in"),
#        dpi = "retina")
# 
# # denses
```

<br>

----


 
# Figure 1

The code below generates Figure 1 from the paper associated with the text on page X.
  
```{r densityrevision, eval = TRUE}

netsize_l <- data.frame(df[, c("netsover6")])
names(netsize_l) <- c("value")
round(summary(df$netsover6)[3], digits = 0)
round(summary(df$netsover6)[4], digits = 0)
IQR(df$netsover6)

dens1 <- ggplot(netsize_l, aes(x=value)) + 
  geom_density(alpha = .2, size = 0.5) + 
  theme_minimal() + 
  theme(legend.position = "none") +
  labs(x = "Network size estimate", y = "Density") +
  scale_color_manual(values = c("#56B4E9")) +
    geom_vline(xintercept=488, color = "darkgrey", linetype = 2) +
  geom_vline(xintercept=568, color = "darkgrey", linetype = 2) +
  annotate("text", x = 630, y = 0.0004, color = "black", angle = 90, 
           label = "Mean = 568") +
  annotate("text", x = 360, y = 0.0004, color = "black", angle = 90, 
           label = "Median = 488") +
  xlim(c(0, 2500)) 

ggsave("output/densities1_revision.pdf", plot = dens1, device = "pdf",
       scale = 1, width = 5, height = 4, units = c("in"),
       dpi = "retina")
ggsave("output/densities1_revision.jpeg", plot = dens1, device = "jpeg",
       scale = 1, width = 5, height = 4, units = c("in"),
       dpi = "retina")


cor(df[, c("netsover1", "netsover2", "netsover4", "netsover5", "netsover6")])
# 
# # VIZ of netsize scenarios
# netsize_l <- gather(df[, c("netsover1", "netsover2", "netsover4", "netsover5", "netsover6")])
# round(summary(df$netsover6)[3], digits = 0)
# round(summary(df$netsover6)[4], digits = 0)
# 
# dens2 <- ggplot(netsize_l, aes(x=value, color = key)) +
#   geom_density(alpha = .2, size = 0.5) +
#   theme_minimal() +
#   theme(legend.position = "none") +
#   labs(x = "Network size estimate", y = "Density") +
#   scale_color_manual(values = c("grey", "grey", "grey", "grey", "#56B4E9")) +
#     geom_vline(xintercept=488, color = "darkgrey", linetype = 2) +
#   geom_vline(xintercept=568, color = "darkgrey", linetype = 2) +
#   annotate("text", x = 560, y = 0.0008, color = "black", angle = 90,
#            label = "Mean = 568") +
#   annotate("text", x = 400, y = 0.0008, color = "black", angle = 90,
#            label = "Median = 488") +
#   xlim(c(0, 2500))
# # save
# ggsave("output/densities2_revision.pdf", plot = dens1, device = "pdf",
#        scale = 1, width = 5, height = 4, units = c("in"),
#        dpi = "retina")
# dens2





```


# Distributional fit

We now fit a "complementary cumulative degree distribution" and check how it fits.


```{r ccdf, eval = TRUE}

#df$netsize <- round(rowSums(df[,c(14:185)]) / length(14:185), 0)
df <- df[!df$netsover6 > 2500, ]
degrees <- round(df$netsover6, 0)

degr_pl <- displ$new(degrees) # create a discrete powerlaw distribution object 

# with estimating lower threshold:
degr_ln_xmin <- dislnorm$new(degrees)
est = estimate_xmin(degr_ln_xmin)
degr_ln_xmin$setXmin(est)

# without lower threshold:
degr_ln_noxmin <-fitdist(degrees, "lnorm")

# Goodness of fit for lognorm, no xmin:
gof_degr_ln_noxmin <- gofstat(degr_ln_noxmin, fitnames = "Log-normal", discrete = TRUE)
gof_degr_ln_noxmin

x_seq <- seq(0, max(degrees), length = 500)

# For plotting the no x_min version of LN
ln_fit_ccdf <- 1 - plnorm(x_seq, meanlog = degr_ln_noxmin$estimate[1], sdlog =  degr_ln_noxmin$estimate[2])

# capture data that make the plot
pd1 <- plot(degr_pl, draw = F) # plot data 1, empirical distr
pd2 <- lines(degr_ln_xmin, draw = F) # plot data 2 no xmin
pd3 <- data.frame(cbind(x_seq, ln_fit_ccdf)) # plot data 3 with xmin

# now render it a ggplot
ccdfplot <- ggplot() + 
              geom_point(data=pd1, aes(x = x, y = y), shape=1) + 
              #geom_line(data=pd2, aes(x = x, y = y, color = "Log-normal fit, w/ x_min")) +
              geom_line(data=pd3, aes(x = x_seq, y = ln_fit_ccdf, color = "Log-normal fit, no x_min")) +
              labs(x="Extended network size", y="CCDF") + 
              scale_color_manual(name = "Fit", values = c(#"Log-normal fit, w/ x_min" = "green", 
                                                          "Log-normal fit, no x_min" = "purple")) +
              theme_bw() +
              theme(legend.position = c(0.22, 0.2)) +
              scale_x_log10(breaks = c(50,100,200,500,1000,2000), limits = c(min(pd1$x), max(pd1$x))) + scale_y_log10() +
              ggtitle("CCDF plot of average degree")

# save
ggsave("output/ccdfplot.pdf", plot = ccdfplot, device = "pdf",
       scale = 1, width = 6, height = 4, units = c("in"),
       dpi = "retina")
ggsave("output/ccdfplot.jpeg", plot = ccdfplot, device = "jpeg",
       scale = 1, width = 6, height = 4, units = c("in"),
       dpi = "retina")

# tryout <- plot_grid(dens2, dens1, ccdfplot, nrow = 1)
# 
# ggsave("output/comps_combined.pdf", plot = tryout, device = "pdf",
#        scale = 1, width = 16, height = 5, units = c("in"),
#        dpi = "retina")

### original plot in Base r
# # Plotting two versions of LN
# plot(degr_pl, xlab="Number of contacts", ylab="CCDF")
# lines(degr_ln_xmin, col = "green", lwd = 2)
# lines(x_seq, ln_fit_ccdf, col = "purple", lwd = 2)
# legend(x = "bottomleft",
#        legend = c("Log-normal fit, w/ x_min",
#                   "Log-normal fit, no x_min" 
#                   ),
#        col = c("green","purple"),
#        lwd = 2)
ccdfplot
```



```{r popvsave, eval = TRUE}

dfans <- df[, c("uni", "hbo", "mbo", "dochterzoon", "tweeling", "corona", 
              "elauto", "scooter", "vegan", 
              "Sophie", "Julia","Sanne","Lisa","Laura","Maria","Linda","Johanna","Monique","Ester",
              "Anna","Elisabeth","Cornelia","Wilhelmina","Amira","Samira","Sara","Daan","Sem",
              "Thomas","Max","Kevin","Johannes","Dennis","Jeroen","Jan","Marcel","Cornelis","Hendrik","Petrus",
              "Willem","Ali","Mohammed","Noor")] # take mean of each of the cats
means <- colMeans(dfans)





ref1 <- c(84957, 75214, 145600, 168066, 2500, 1558549)
# ORDER:  uni, hbo, mbo, dochter/zoon, tweeling, corona


ref2 <- c(273259, 460618, 261000)
# ORDER:   elecauto, scooter, vegan,

ref3 <- c(15276,16350,27394,21200,25681,334502,29955,266522,39481, 2692,
          136296,110231,112807,98208,1386,2186,11640,22704,13276,
          40543,17024,23167,307032,36411,49182,186746,35973,134956,118610,86500,
          102296,4213,5003,4517)
# ORDER:   Sophie, Julia,Sanne,Lisa,Laura,Maria,Linda,Johanna,Monique,Ester,
#             Anna,Elisabeth,Cornelia,Wilhelmina,Amira,Samira,Sara,Daan,Sem,
#             Thomas,Max,Kevin,Johannes,Dennis,Jeroen,Jan,Marcel,Cornelis,Hendrik,Petrus,
#             Willem,Ali,Mohammed,Noor

pops <- c(ref1, ref2, ref3)

fig <- data.frame(cbind(means, pops))



cor(fig[!fig$pops > 500000, ])
cor(fig[!fig$pops > 200000, ])
cor(fig[fig$pops < 50000, ])
cor(fig[fig$pops < 100000, ])


ex3 <- ggplot(fig, aes(x=pops, y=means)) + 
  geom_point() + 
  xlim(x= c(0, 50000)) + 
  ylim(c(0,2.5)) + 
  geom_smooth(method='lm') +
  geom_text(label = rownames(fig),
            nudge_y = 0.1) +
  theme_minimal() +
  xlab("Population size") +
  ylab("Average number mentioned") +
  sm_statCorr() + 
  ggtitle("C) Population size <50k")
ex3

ex2 <- ggplot(fig, aes(x=pops, y=means)) + 
  geom_point() + 
  xlim(x= c(0, 500000)) + 
  ylim(c(0,2.5)) + 
  geom_smooth(method='lm') +
  geom_text(label = rownames(fig),
            nudge_y = 0.1) +
  theme_minimal() +
  xlab("Population size") +
  ylab("Average number mentioned") +
  sm_statCorr()+ 
  ggtitle("B) Population size <500k")
ex2

ex1 <- ggplot(fig, aes(x=pops, y=means)) + 
  geom_point() + 
  geom_smooth(method='lm') +
  geom_text(label = rownames(fig),
            nudge_y = 0.1) +
  theme_minimal() +
  xlab("Population size") +
  ylab("Average number mentioned") +
  sm_statCorr() + 
  ggtitle("A) All population sizes")
ex1

ex <- cowplot::plot_grid(ex1, ex2, ex3, nrow = 1)

ggsave("output/popsizes.pdf", plot = ex, device = "pdf",
       scale = 1, width = 16, height = 5, units = c("in"),
       dpi = "retina")
ggsave("output/popsizes.jpeg", plot = ex, device = "jpeg",
       scale = 1, width = 16, height = 5, units = c("in"),
       dpi = "retina")
ex
```






