---
title: "Network size of the Dutch"
output:
  html_document:
    toc: false
    number_sections: false
author: "Beate Volker, Bas Hofstra, Rense Corten, Frank van Tubergen"
bibliography: references.bib

---

```{r, install remotely, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
# install.packages("remotes")
#remotes::install_github("rlesur/klippy")
``` 

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```




----



# Network size of the Dutch
Last compiled on `r format(Sys.time(), '%B, %Y')`  


<br>

----

## Replication website

This is the replication website for the paper entitled: "How large is your network? Estimating Network Size of Residents in the Netherlands." By Beate Volker, Bas Hofstra, Rense Corten, and Frank van Tubergen. 


## How to use this website

The GitHub repository to this website is found [here](https://github.com/bhofstra/netsize_dutch). Each page captures a different aspect to the research paper. 

## Acknowledgements

We acknowledge Rob Franken for his LabJournal example on which we based this replication website, found [here](https://github.com/robfranken/LabJournal).  


----






