
---
title: "Inferential analyses"
#bibliography: references.bib
author: "Bas Hofstra"
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r format(Sys.time(), '%B, %Y')`

<br>

----

This is the code with which we run our inferential analyses.

<br>

----

# Initatiating R environment

Start out with a custom function to load a set of required packages.
  
```{r pack, eval=TRUE}
# packages and read data
rm(list = ls())

# (c) Jochem Tolsma
fpackage.check <- function(packages) {
  lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  })
}
packages = c("haven", "coda", "matrixStats", "parallel", "MASS", "doParallel", "dplyr", "cowplot", 
             "tidyverse", "naniar", "dotwhisker" ,"gt", "reshape2", "VGAM", "expss", "Hmisc", "MASS", "sjPlot")
fpackage.check(packages)
rm(packages)
load("data/dutch_netsize_analyses_revision.rda")
```

<br>

----

# Linear regression models

We run a series of regression models of logged acquantanceship network size (as it is right-skewed) on our independent and control variables. We thus run 172 different regression models (all scenarios).

```{r regs, eval = TRUE}
# 
# 
# ########################
# # MODELS
# #######################
# 
# # # age in categories, eval = false in "independent_variables.rmd" fix later
# df$agecat[df$agecat == "18-30"] <- 1
# df$agecat[df$agecat == "31-45"] <- 2
# df$agecat[df$agecat == "46-65"] <- 3
# df$agecat[df$agecat == ">65"] <- 4
# 
# table(df$agecat)
# class(df$agecat)
# 
# 
# # some final data handling to assign the correct reference categories
# df$opl <-relevel(df$opl, ref = 1)
# df$migr3 <-relevel(as.factor(df$migr3), ref = 1)
# df$income <-relevel(as.factor(df$income), ref = 2)
# df$agecat <-relevel(as.factor(df$agecat), ref = 4)
# 
# 
# 
# 
#  # 172 regressions with different netsizes. We then look at distributions of coefficients across those.
# # note the log10 for network size and how we don't take into account extreme network size.
# modellog <- list()
# for (i in 14:185) {
#     
#     df[,c(i)] <- round(df[, c(i)], 0)
#     
#     modellog[[i]] <- lm(log10(df[!df[[i]]>5000, c(i)]) ~ 
#                           
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr3) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                           
#                         data = df[!df[[i]]>5000,])
#     
# }
# modellog[sapply(modellog, is.null)] <- NULL
# summary(modellog[[1]])
```

<br>

----

# Visualization of regression models

We then render all those models into one figure where we can show statistical signifance and that results do not vary over all those scenarios. This matches Figure 3 in the paper.


```{r regviz, eval = TRUE}

# #--------------------------------------------------------------------------------
# # Viz of results
# 
# # main effects
# four_brackets <- list(
#   
#   c("Dutch maj=ref", "West backgr", "non-West backgr"),
#   
#   c(">65=ref", "18-30", "46-65"),
#   
#   c(">modal inc=ref", "<=modal income", "Unkn income"),
#   
#   c("Educ tert high=ref", "Educ prim/sec", "Educ tert low"))
# 
#  logmodel <- {dwplot(modellog,
#          dot_args = list(color = "black", size = 0.75, shape = 1), # color for the dot
#          whisker_args = list(size = 0.25, color = "darkgrey", alpha = 0.5),   # color for the whisker
#          vline = ggplot2::geom_vline(xintercept = 0,  # put vline _behind_ coefs; see https://github.com/fsolt/dotwhisker/issues/84
#                                      colour = "grey60",
#                                      linetype = 2,
#                                      linewidth = 1)) %>%                                     # make model variable
#     relabel_predictors(c(
#       "as.factor(work)1" = "Working",
#       
#       "hhsize" = "Household size",
#       
#       "as.factor(migr3)2" = "West backgr",
#       "as.factor(migr3)3" = "non-West backgr",
#       
#       "as.factor(agecat)1" = "18-30",
#       "as.factor(agecat)2" = "31-45",
#       "as.factor(agecat)3" = "46-65",
#       
#       "as.factor(income)1" = "<=modal income",
#       "as.factor(income)3" = "Unkn income",
#       
#       "worthhouse" = "House value",
# 
#       "as.factor(woman)1" = "Women",
#       
#       "as.factor(opl)1" = "Educ prim/sec",
#       "as.factor(opl)2" = "Educ tert low"
# 
#     )) + theme(legend.position = "none",
#                axis.text = element_text(color = "grey")) + 
#     theme_minimal() +
#     xlab("B on Log(network size count)")}  %>%
#   add_brackets(four_brackets, fontSize = .6)
# 
#  ggsave("output/models.pdf", plot = logmodel, device = "pdf",
#        scale = 1, width = 6, height = 5, units = c("in"),
#        dpi = "retina")
# 
#  # Figure 3 in the paper
# logmodel
#  

```


<br>

----

# Regression table

We also average all 172 network size scenarios within respondents and run one single regression model so as to generate a table with coefficients. This matches Table 3 in the paper.


```{r regtable, eval = TRUE}
# 
# df$netsize <- round(rowSums(df[,c(14:185)]) / length(14:185), 0)
# 
# modell <- lm(log10(netsize) ~ 
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr3) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                         data = df[!df[["netsize"]]>5000,])
# 
# modellin <- lm(netsize ~ 
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr3) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                         data = df[!df[["netsize"]]>5000,])
#   
#   
# fpackage.check("sjPlot")
# 
# #table 3 in the paper
# tab_model(modell, modellin, show.se = TRUE)
```

```{r regrobust, eval = TRUE}
# # robustness, migration two cats, and income no missings lead to similar results
# summary(x <-lm(log10(netsize) ~ 
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                         data = df[!df[["netsize"]]>5000,]))
# 
# summary(x <-lm(log10(netsize) ~ 
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr3) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                         data = df[!df[["netsize"]]>5000 & df$income != 3,]))


```

----

# Revisions Regression table


From here onwards, we show our new results based on the revisions of this paper.

```{r regrevision_size, eval = TRUE}

# # age in categories, eval = false in "independent_variables.rmd" fix later
df$agecat[df$agecat == "18-30"] <- 1
df$agecat[df$agecat == "31-45"] <- 2
df$agecat[df$agecat == "46-65"] <- 3
df$agecat[df$agecat == ">65"] <- 4

table(df$agecat)
class(df$agecat)


# some final data handling to assign the correct reference categories
#df$opl2 <-relevel(as.factor(df$opl2), ref = 0)

df$agecat<- as.character(df$agecat)
df$income<- as.character(df$income)
df$income <-relevel(as.factor(df$income), ref = 2)
df$agecat <-relevel(as.factor(df$agecat), ref = 4)

table(df$opl2)

modell2 <- lm(log(netsover4+1) ~ 
                          as.factor(work) + #h1
                           
                          hhsize + #H2

                          as.factor(agecat) + #h3
                          
                          as.factor(income) + #h4
                          
                          worthhouse + #H4
                          
                          as.factor(woman) + #h5
                          
                          as.factor(opl2), #h6
                        data = df[!df[["netsover4"]]>5000,])
summary(modell2)


modellnb2 <- glm.nb(netsover4 ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl2),
                        data = df[!df[["netsover4"]]>5000,],
                        init.theta = 1.032713156, link = log)
summary(modellnb2)

fpackage.check("sjPlot")

# table 3 paper:
tab_model(modell2, modellnb2, show.se = TRUE)
```

```{r regrevision_gen, eval = TRUE}
#gender
modelwoman <- lm(numwoman ~ 
                          as.factor(work) +
                           
                          hhsize + 
                   
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse + 
                          
                          as.factor(woman) + #h7
                          
                          as.factor(opl2), 
                        data = df)
summary(modelwoman)


modelman <- lm(numman ~ 
                         as.factor(work) +
                           
                          hhsize + 
                   
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse + 
                          
                          as.factor(woman) + #h7
                          
                          as.factor(opl2), 
                        data = df)
summary(modelman)


modelsamegen <- lm(samegender ~ 
                     
                         as.factor(work) +
                           
                          hhsize + 
                   
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse + 
                          
                          as.factor(woman) + #h7
                          
                          as.factor(opl2), 
                        data = df)
summary(modelsamegen)

tab_model(modelwoman, modelman, modelsamegen, show.se = TRUE)
```


```{r regrevision_educ, eval = TRUE}

# line educ independent variable up to educ homogeneity dependent

#educ
modeleduch <- lm(numeduchigh ~ 
                          as.factor(work) +
                           
                          hhsize + 
                   
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse + 
                          
                          as.factor(woman) + 
                          
                          as.factor(opl2), # h8
                        data = df)
summary(modeleduch)

modeleducl <- lm(numeduclow ~ 
                          as.factor(work) +
                           
                          hhsize + 
                   
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse + 
                          
                          as.factor(woman) + 
                          
                          as.factor(opl2), # h8
                        data = df)
summary(modeleducl)

modelsameeduc <- lm(sameeduc ~ 
                          as.factor(work) +
                           
                          hhsize + 
                   
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse + 
                          
                          as.factor(woman) + 
                          
                          as.factor(opl2), # h8
                        data = df)
summary(modelsameeduc)

tab_model(modeleduch, modeleducl, modelsameeduc, show.se = TRUE)

```

Hypothesis 9: network size by homogeneity.

```{r regrevision_h9, eval = TRUE}

genh9 <- lm(samegender ~ 
                          as.factor(work) +
                           
                          hhsize + 
                   
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse + 
                          
                          as.factor(woman) + 
                          
                          as.factor(opl2) +
                          
                          log(netsover4+1), #H9
                        data = df)


educh9 <- lm(sameeduc ~ 
                          as.factor(work) +
                           
                          hhsize + 
                   
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse + 
                          
                          as.factor(woman) + 
                          
                          as.factor(opl2) +
                          
                          log(netsover4+1), #H9
                        data = df)

summary(genh9)
summary(educh9)
tab_model(genh9, educh9, show.se = TRUE)

```

Ethnic background is dropped from the paper.

```{r regrevision_eth, eval = TRUE}
# 
# #gender
# modeldutch <- lm(numdutch ~ 
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr3) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                           data = df)
# #summary(modeldutch)
# 
# 
# modelnodutch <- lm(numnodutch ~ 
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr3) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                         data = df)
# #summary(modelnodutch)
# 
# 
# modelsameethnic1 <- lm(sameethnic1 ~ 
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr3) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                         data = df)
# #summary(modelsameethnic1)
# 
# tab_model(modeldutch, modelnodutch, modelsameethnic1, show.se = TRUE)
```


# Revisions Robustness Regression table

```{r regrevision_robust, eval = TRUE}


# robustness four other estimation scenarios: mostly qualitatively similar
r1 <- lm(log(netsover1+1) ~ 
                          as.factor(work) + #h1
                           
                          hhsize + #H2

                          as.factor(agecat) + #h3
                          
                          as.factor(income) + #h4
                          
                          worthhouse + #H4
                          
                          as.factor(woman) + #h5
                          
                          as.factor(opl2), #h6
                        data = df[!df[["netsover1"]]>5000,])


r2 <- lm(log(netsover2+1) ~ 
                          as.factor(work) + #h1
                           
                          hhsize + #H2

                          as.factor(agecat) + #h3
                          
                          as.factor(income) + #h4
                          
                          worthhouse + #H4
                          
                          as.factor(woman) + #h5
                          
                          as.factor(opl2), #h6
                        data = df[!df[["netsover2"]]>5000,])

r3 <- lm(log(netsover3+1) ~ 
                          as.factor(work) + #h1
                           
                          hhsize + #H2

                          as.factor(agecat) + #h3
                          
                          as.factor(income) + #h4
                          
                          worthhouse + #H4
                          
                          as.factor(woman) + #h5
                          
                          as.factor(opl2), #h6
                        data = df[!df[["netsover3"]]>5000,])

r5 <- lm(log(netsover5+1) ~ 
                          as.factor(work) + #h1
                           
                          hhsize + #H2

                          as.factor(agecat) + #h3
                          
                          as.factor(income) + #h4
                          
                          worthhouse + #H4
                          
                          as.factor(woman) + #h5
                          
                          as.factor(opl2), #h6
                        data = df[!df[["netsover5"]]>5000,])

# robustness for scenarios
summary(r1)
summary(r2)
summary(r3)
summary(r5)



```