
---
title: "Inferential analyses"
#bibliography: references.bib
author: "Bas Hofstra"
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r format(Sys.time(), '%B, %Y')`

<br>

----

This is the code with which we run our inferential analyses.

<br>

----

# Initatiating R environment

Start out with a custom function to load a set of required packages.
  
```{r pack, eval=TRUE}
# packages and read data
rm(list = ls())

# (c) Jochem Tolsma
fpackage.check <- function(packages) {
  lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  })
}
packages = c("haven", "coda", "matrixStats", "parallel", "MASS", "doParallel", "dplyr", "cowplot", 
             "tidyverse", "naniar", "dotwhisker" ,"gt", "reshape2", "VGAM", "expss", "Hmisc")
fpackage.check(packages)
rm(packages)
load("data/dutch_netsize_analyses.rda")
```

<br>

----

# Linear regression models

We run a series of regression models of logged acquantanceship network size (as it is right-skewed) on our independent and control variables. We thus run 172 different regression models (all scenarios).

```{r regs, eval = TRUE}


########################
# MODELS
#######################


# some final data handling to assign the correct reference categories
df$opl <-relevel(df$opl, ref = 1)
df$migr3 <-relevel(as.factor(df$migr3), ref = 1)
df$income <-relevel(as.factor(df$income), ref = 2)

 # 172 regressions with different netsizes. We then look at distributions of coefficients across those.
# note the log10 for network size and how we don't take into account extreme network size.
modellog <- list()
for (i in 14:185) {
    
    df[,c(i)] <- round(df[, c(i)], 0)
    
    modellog[[i]] <- lm(log10(df[!df[[i]]>5000, c(i)]) ~ 
                          as.factor(work) + 
                          hhsize + 
                          as.factor(income) + 
                          leeftijd10 + 
                          as.factor(migr3) + 
                          as.factor(woman) + 
                          as.factor(opl) + 
                          neighdens + 
                          worthhouse,
                        data = df[!df[[i]]>5000,])
    
}
modellog[sapply(modellog, is.null)] <- NULL
```

<br>

----

# Visualization of regression models

We then render all those models into one figure where we can show statistical signifance and that results do not vary over all those scenarios. This matches Figure 3 in the paper.


```{r regviz, eval = TRUE}
#--------------------------------------------------------------------------------
# Viz of results

# main effects
three_brackets <- list(
  c(">modal inc=ref", "<=modal income", "Unkn income"),
  c("Dutch maj=ref", "West backgr", "non-West backgr"),
  c("Educ tert high=ref", "Educ prim/sec", "Educ tert low"))

 logmodel <- {dwplot(modellog,
         dot_args = list(color = "black", size = 0.75, shape = 1), # color for the dot
         whisker_args = list(size = 0.25, color = "darkgrey", alpha = 0.5),   # color for the whisker
         vline = ggplot2::geom_vline(xintercept = 0,  # put vline _behind_ coefs; see https://github.com/fsolt/dotwhisker/issues/84
                                     colour = "grey60",
                                     linetype = 2,
                                     size = 1)) %>%                                     # make model variable
    relabel_predictors(c(
      "as.factor(work)1" = "Working",
      "hhsize" = "Household size",
      "as.factor(income)1" = "<=modal income",
      "as.factor(income)3" = "Unkn income",
      "leeftijd10" = "Age/10",
      "as.factor(migr3)2" = "West backgr",
      "as.factor(migr3)3" = "non-West backgr",
      "as.factor(woman)1" = "Women",
      "as.factor(opl)1" = "Educ prim/sec",
      "as.factor(opl)2" = "Educ tert low",
      "neighdens" = "Neigh. density",
      "worthhouse" = "WOZ-values"
    )) + theme(legend.position = "none",
               axis.text = element_text(color = "grey")) + 
    theme_minimal() +
    xlab("B on Log(network size count)")}  %>%
  add_brackets(three_brackets, fontSize = .6)

 ggsave("output/models.pdf", plot = logmodel, device = "pdf",
       scale = 1, width = 6, height = 5, units = c("in"),
       dpi = "retina")

 # Figure 3 in the paper
logmodel
 

```


<br>

----

# Regression table

We also average all 172 network size scenarios within respondents and run one single regression model so as to generate a table with coefficients. This matches Table 3 in the paper.


```{r regtable, eval = TRUE}

df$netsize <- round(rowSums(df[,c(14:185)]) / length(14:185), 0)

modell <- lm(log10(netsize) ~ 
                          as.factor(work) + 
                          hhsize + 
                          as.factor(income) + 
                          leeftijd10 + 
                          as.factor(migr3) + 
                          as.factor(woman) + 
                          as.factor(opl) + 
                          neighdens + 
                          worthhouse,
                        data = df[!df[["netsize"]]>5000,])

modellin <- lm(netsize ~ 
                          as.factor(work) + 
                          hhsize + 
                          as.factor(income) + 
                          leeftijd10 + 
                          as.factor(migr3) + 
                          as.factor(woman) + 
                          as.factor(opl) + 
                          neighdens + 
                          worthhouse,
                        data = df[!df[["netsize"]]>5000,])
  
  
fpackage.check("sjPlot")

#table 3 in the paper
tab_model(modell, modellin, show.se = TRUE)
```

