
---
title: "Inferential analyses"
#bibliography: references.bib
author: "Bas Hofstra"
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r format(Sys.time(), '%B, %Y')`

<br>

----

This is the code with which we run our inferential analyses.

<br>

----

# Initatiating R environment

Start out with a custom function to load a set of required packages.
  
```{r pack, eval=TRUE}
# packages and read data
rm(list = ls())

# (c) Jochem Tolsma
fpackage.check <- function(packages) {
  lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  })
}
packages = c("haven", "coda", "matrixStats", "parallel", "MASS", "doParallel", "dplyr", "cowplot", 
             "tidyverse", "naniar", "dotwhisker" ,"gt", "reshape2", "VGAM", "expss", "Hmisc", "MASS", "sjPlot")
fpackage.check(packages)
rm(packages)
load("data/dutch_netsize_analyses.rda")
```

<br>

----

# Linear regression models

We run a series of regression models of logged acquantanceship network size (as it is right-skewed) on our independent and control variables. We thus run 172 different regression models (all scenarios).

```{r regs, eval = TRUE}
# 
# 
# ########################
# # MODELS
# #######################
# 
# # # age in categories, eval = false in "independent_variables.rmd" fix later
# df$agecat[df$agecat == "18-30"] <- 1
# df$agecat[df$agecat == "31-45"] <- 2
# df$agecat[df$agecat == "46-65"] <- 3
# df$agecat[df$agecat == ">65"] <- 4
# 
# table(df$agecat)
# class(df$agecat)
# 
# 
# # some final data handling to assign the correct reference categories
# df$opl <-relevel(df$opl, ref = 1)
# df$migr3 <-relevel(as.factor(df$migr3), ref = 1)
# df$income <-relevel(as.factor(df$income), ref = 2)
# df$agecat <-relevel(as.factor(df$agecat), ref = 4)
# 
# 
# 
# 
#  # 172 regressions with different netsizes. We then look at distributions of coefficients across those.
# # note the log10 for network size and how we don't take into account extreme network size.
# modellog <- list()
# for (i in 14:185) {
#     
#     df[,c(i)] <- round(df[, c(i)], 0)
#     
#     modellog[[i]] <- lm(log10(df[!df[[i]]>5000, c(i)]) ~ 
#                           
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr3) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                           
#                         data = df[!df[[i]]>5000,])
#     
# }
# modellog[sapply(modellog, is.null)] <- NULL
# summary(modellog[[1]])
```

<br>

----

# Visualization of regression models

We then render all those models into one figure where we can show statistical signifance and that results do not vary over all those scenarios. This matches Figure 3 in the paper.


```{r regviz, eval = TRUE}

# #--------------------------------------------------------------------------------
# # Viz of results
# 
# # main effects
# four_brackets <- list(
#   
#   c("Dutch maj=ref", "West backgr", "non-West backgr"),
#   
#   c(">65=ref", "18-30", "46-65"),
#   
#   c(">modal inc=ref", "<=modal income", "Unkn income"),
#   
#   c("Educ tert high=ref", "Educ prim/sec", "Educ tert low"))
# 
#  logmodel <- {dwplot(modellog,
#          dot_args = list(color = "black", size = 0.75, shape = 1), # color for the dot
#          whisker_args = list(size = 0.25, color = "darkgrey", alpha = 0.5),   # color for the whisker
#          vline = ggplot2::geom_vline(xintercept = 0,  # put vline _behind_ coefs; see https://github.com/fsolt/dotwhisker/issues/84
#                                      colour = "grey60",
#                                      linetype = 2,
#                                      linewidth = 1)) %>%                                     # make model variable
#     relabel_predictors(c(
#       "as.factor(work)1" = "Working",
#       
#       "hhsize" = "Household size",
#       
#       "as.factor(migr3)2" = "West backgr",
#       "as.factor(migr3)3" = "non-West backgr",
#       
#       "as.factor(agecat)1" = "18-30",
#       "as.factor(agecat)2" = "31-45",
#       "as.factor(agecat)3" = "46-65",
#       
#       "as.factor(income)1" = "<=modal income",
#       "as.factor(income)3" = "Unkn income",
#       
#       "worthhouse" = "House value",
# 
#       "as.factor(woman)1" = "Women",
#       
#       "as.factor(opl)1" = "Educ prim/sec",
#       "as.factor(opl)2" = "Educ tert low"
# 
#     )) + theme(legend.position = "none",
#                axis.text = element_text(color = "grey")) + 
#     theme_minimal() +
#     xlab("B on Log(network size count)")}  %>%
#   add_brackets(four_brackets, fontSize = .6)
# 
#  ggsave("output/models.pdf", plot = logmodel, device = "pdf",
#        scale = 1, width = 6, height = 5, units = c("in"),
#        dpi = "retina")
# 
#  # Figure 3 in the paper
# logmodel
#  

```


<br>

----

# Regression table

We also average all 172 network size scenarios within respondents and run one single regression model so as to generate a table with coefficients. This matches Table 3 in the paper.


```{r regtable, eval = TRUE}
# 
# df$netsize <- round(rowSums(df[,c(14:185)]) / length(14:185), 0)
# 
# modell <- lm(log10(netsize) ~ 
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr3) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                         data = df[!df[["netsize"]]>5000,])
# 
# modellin <- lm(netsize ~ 
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr3) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                         data = df[!df[["netsize"]]>5000,])
#   
#   
# fpackage.check("sjPlot")
# 
# #table 3 in the paper
# tab_model(modell, modellin, show.se = TRUE)
```

```{r regrobust, eval = TRUE}
# # robustness, migration two cats, and income no missings lead to similar results
# summary(x <-lm(log10(netsize) ~ 
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                         data = df[!df[["netsize"]]>5000,]))
# 
# summary(x <-lm(log10(netsize) ~ 
#                           as.factor(work) + 
#                           
#                           hhsize + 
#                           
#                           as.factor(migr3) + 
#                           
#                           as.factor(agecat) + 
#                           
#                           as.factor(income) + 
#                           
#                           worthhouse +
#                           
#                           as.factor(woman) + 
#                           
#                           as.factor(opl),
#                         data = df[!df[["netsize"]]>5000 & df$income != 3,]))


```

```{r regrevision_size, eval = TRUE}


# # age in categories, eval = false in "independent_variables.rmd" fix later
df$agecat[df$agecat == "18-30"] <- 1
df$agecat[df$agecat == "31-45"] <- 2
df$agecat[df$agecat == "46-65"] <- 3
df$agecat[df$agecat == ">65"] <- 4

table(df$agecat)
class(df$agecat)


# some final data handling to assign the correct reference categories
df$opl <-relevel(df$opl, ref = 1)
df$migr3 <-relevel(as.factor(df$migr3), ref = 1)
df$income <-relevel(as.factor(df$income), ref = 2)
df$agecat <-relevel(as.factor(df$agecat), ref = 4)



modell2 <- lm(log(netsover2+1) ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df[!df[["netsover2"]]>5000,])
summary(modell2)


modellnb2 <- glm.nb(netsover2 ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df[!df[["netsover2"]]>5000,],
                        init.theta = 1.032713156, link = log)
summary(modellnb2)

fpackage.check("sjPlot")

tab_model(modell2, modellnb2, show.se = TRUE)
```

```{r regrevision_gen, eval = TRUE}
#gender
modelwoman <- lm(numwoman ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df)
# summary(modelwoman)


modelman <- lm(numman ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df)
# summary(modelman)


modelsamegen <- lm(samegender ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df)
# summary(modelsamegen)

tab_model(modelwoman, modelman, modelsamegen, show.se = TRUE)
```


```{r regrevision_educ, eval = TRUE}

#gender
modelmbo <- lm(mbo ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df)
# summary(modelmbo)


modelhbo <- lm(hbo ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df)
# summary(modelhbo)


modeluni <- lm(uni ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df)
# summary(modeluni)


modelsameeduc <- lm(sameeduc ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df)
# summary(modelsameeduc)

tab_model(modelmbo, modelhbo, modeluni, modelsameeduc, show.se = TRUE)

```

```{r regrevision_eth, eval = TRUE}

#gender
modeldutch <- lm(numdutch ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df)
#summary(modeldutch)


modelnodutch <- lm(numnodutch ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df)
#summary(modelnodutch)


modelsameethnic1 <- lm(sameethnic1 ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df)
#summary(modelsameethnic1)

tab_model(modeldutch, modelnodutch, modelsameethnic1, show.se = TRUE)
```


```{r regrevision_robust, eval = TRUE}

modell1 <- lm(log(netsover1+1) ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df[!df[["netsover1"]]>5000,])
summary(modell1)


modelwoman <- lm(numwoman) ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df)
modelwoman <- lm(numwoman) ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df)
summary(modell1)



modell3 <- lm(log(netsover3+1) ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df[!df[["netsover3"]]>5000,])
summary(modell3)

modell4 <- lm(log(netsover4+1) ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df[!df[["netsover4"]]>5000,])
summary(modell4)

modellnb1 <- glm.nb(netsover1 ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df[!df[["netsover1"]]>5000,],
                        init.theta = 1.032713156, link = log)
summary(modellnb1)


modellnb3 <- glm.nb(netsover3 ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df[!df[["netsover3"]]>5000,],
                        init.theta = 1.032713156, link = log)
summary(modellnb3)


modellnb4 <- glm.nb(netsover4 ~ 
                          as.factor(work) + 
                          
                          hhsize + 
                          
                          as.factor(migr3) + 
                          
                          as.factor(agecat) + 
                          
                          as.factor(income) + 
                          
                          worthhouse +
                          
                          as.factor(woman) + 
                          
                          as.factor(opl),
                        data = df[!df[["netsover4"]]>5000,],
                        init.theta = 1.032713156, link = log)
summary(modellnb4)



```